
\chapter*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}


First and foremost I'd like to thank Matthew Hole, my supervisor. I came 
into this year a little daunted at the prospect of tackling a physics-based 
project with a physics background not extending past ``A Brief History of Time''.
Your patience and willingness to explain concepts to me I'll forever appreciate, 
and your energetic approach to problem solving made me feel both at home and excited 
for the work we have done in this thesis. Thank you for always taking the time to talk to 
me, and for being a fantastic supervisor.

In a similar vein, thank you to the rest of the Plasma Theory and Modelling group for always being 
willing to answers questions when I had them. You guys were definitely my saving grace 
at times, and I only wish I spent more time getting to know you all this year. I should give 
a special shout out to Nick, Sandra, Dean and Josh specifically for being so 
accommodating, helpful, and kind in answering my questions when I would pop around. 

I'd like to thank the other Honours students for helping make this a fantastic 
year. Our little nook of MSI proved very distracting, but in the best possible 
way. As such I would of course expect you to take part responsibility for the late night 
writing sessions at the end of the year - but you also get 
credit for making them enjoyable. Also Joe gets a special shout out purely to spite everyone else.

Last but certainly not least, to Kirsten and Graeme. Thank you for your support these last 
couple years, and the infinite kindness you give to all those around you. However silly it is,
your encouragement genuinely helped motivate me at times, and it was always my pleasure having the 
opportunity to rant to you about whatever problem I was working on at the time, and how insurmountable
it seemed! Yet each time you had confidence it would all work out fine, and then somehow it always would. I can only assume 
this is some wizardry I am not yet privy to, but am nevertheless indebted to your kindness and hospitality.